---
title: "DATASET ANALYSIS"
author: "Elizabeth Edminster"
date: "2024-10-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
df <- read.csv("CleanedDf.csv")
```

## Summary of Data
      Column descriptions

**general info**      
  LIBNAME: Name of library
  STABR: State
  C_LEGBAS: Legal Basis Code
CC–City/County
CI–Municipal Government (city, town, or village)
CO–County/Parish
LD–Library District
MJ–Multi-jurisdictional
NL–Native American Tribal Government
NP–Non-profit Association or Agency
SD–School District
OT–Other
  POPU_LSA: Population of the legal service area (population group that the library caters to)

**locations & employees**
  CENTLIB: Number of central libraries
  BRANLIB: Number of branch libraries
  BKMOB: Number of bookmobiles
  MASTER: Number of full-time librarians with masters degrees in library science
  LIBRARIA: Number of full-time librarians.

**revenue**
  TOTINCM: *Total operating revenue* (DO NOT include revenue for major capital expenditures, contributions to endowments, revenue passed through to another agency (e.g., fines), or funds unspent in the previous fiscal year (e.g., carryover)). This is revenue given by federal, state, & local governments, monetary gifts and donations received in the current year, interest, library fines, fees for library services, etc.
  CAP_REV: *Total capital revenue* (revenue to be used for fixed cost items (funds
received for (a) site acquisition; (b) new buildings; (c) additions to or renovation        of library buildings; (d) furnishings, equipment, and initial collections (print,           nonprint, and electronic) for new buildings, building additions, or building                renovations; (e) computer hardware and software used to support library operations, to      link to networks, or to run information products; (f) new vehicles; and (g) other           one-time major projects.))

**operating expenditures**
  PRMATEXP: Total operating cost for print materials
  ELMATEXP: Total operating cost for electronic materials
  OTHMATEX: Total operating cost for other materials

**capital expenditures**
  CAPITAL: Total capital expenditures

**collection**
  BKVOL: Print book collection (including duplicates)
  EBOOK: EBook licencing collection (include duplicates)
  VIDEO_PH: Physical video units (including duplicates)
  AUDIO_PH: Physical audio units (including duplicates)
  VIDEO_DL: Downloadable video units (including duplicates)
  AUDIO_DL: Downloadable audio units (including duplicates)
 
**metrics**
  VISITS: total persons entering the library per year (estimated or counted)
  REGBOR: total registered users
  ELMATCIR: Total use of Electronic materials
  PHYSCIR: Total use of Physical materials
  KIDCIRCL: Total use of Childrens materials, electronic or not.
  PITUSR: Total use of internet-connected computers

**inter-library relations**
  LOANTO: quantity of library materials provided to other libraries
  LOANFM: quantity of library materials recieved from other libraries

```{r summary}
summary(df)
```

```{r}
head(df)
```

```{r}
ncol(df)
nrow(df)
```
##Binning

In order to compare libraries of different size categories more easily, I decided to bin them (turn them into a categorical variable).

This is the preliminary distribution of the libraries
```{r}
POPU_LSA_BoxPlot <- boxplot(df$POPU_LSA)
POPU_LSA_BoxPlot
```
I binned the libraries into 10 groups of a bit less than 100 libraries each. 
```{r}
df <- df %>%
  mutate(limit = ntile(POPU_LSA, 10)) %>% 
  mutate(limitMarker = case_when(
    limit == 10 ~ "Up90",
    limit == 9 ~ "80to90",
    limit == 8 ~ "70to80",
    limit == 7 ~ "60to70",
    limit == 6 ~ "50to60",
    limit == 5 ~ "40to50",
    limit == 4 ~ "30to40",
    limit == 3 ~ "20to30",
    limit == 2 ~ "10to20",
    limit == 1 ~ "Down10"
  ))
```

The new boxplot of just libraries below percentile 80 is below
```{r}
New_POPU_LSA_BoxPlot <- df %>% filter(!(limitMarker %in% c("Up90", "80to90")))

New_POPU_LSA_BoxPlot <- boxplot(New_POPU_LSA_BoxPlot$POPU_LSA)

New_POPU_LSA_BoxPlot
```
Another way of looking at it is with a histogram. 

All libraries:
```{r}
POPU_LSA_Histogram <- hist(df$POPU_LSA)
POPU_LSA_Histogram
```

Only libraries below 80th percentile:  
```{r}
New_POPU_LSA_Histogram <- df %>% filter(!(limitMarker %in% c("Up90", "80to90")))
New_POPU_LSA_Histogram <- hist(New_POPU_LSA_Histogram$POPU_LSA)
```
The purpose of this binning is to
  a. Eliminate outliers that may skew results
  b. Draw conclusions on libraries that perform with relatively similar business models (the NYC public library is not going to have the same business model as your local library, but your local library may have a similar business model to a library 3 towns away.)
  
Also, I chose to bin the states in order to provide another level of categorization.
```{r}
df <- df %>%
  mutate(Region = case_when(
    STABR %in% c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA", "DE", "MD", "DC") ~ "New England & Mid-Atlantic",
    STABR %in% c("VA", "WV", "KY", "NC", "SC", "GA", "FL", "AL", "MS", "TN") ~ "Southeast",
    STABR %in% c("OH", "MI", "IN", "IL", "WI", "MN") ~ "Great Lakes",
    STABR %in% c("IA", "MO", "ND", "SD", "NE", "KS") ~ "Midwest & Plains",
    STABR %in% c("TX", "OK", "AR", "LA", "NM", "AZ") ~ "Southwest",
    STABR %in% c("CA", "NV", "OR", "WA", "ID", "UT", "CO", "MT", "WY", "AK", "HI") ~ "West",
    TRUE ~ "Other"
  ))
```

```{r}
State_boxplot <- boxplot(df$EBOOK ~ df$Region)
```


## Trends


