---
title: "DATASET PREP OIM454"
author: "Elizabeth Edminster"
date: "2024-09-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#if you are running this from your pc, adjust the filepath
df <- read.csv("C:/Users/lized/Downloads/pls_fy2022_csv/PLS_FY2022 PUD_CSV/PLS_FY22_AE_pud22i.csv")
```
## Quantitative variable summary
The following results show the summary statistics of all quantitative variables (note: certain variables like zip show summary statistics, but these are not relevant for analysis in this way)
Number of records(before cleaning) = 9,248
Number of columns(before cleaning) = 192
```{r summarystats}
q4 <- data.frame(unclass(summary(df)), check.names = FALSE, stringsAsFactors = FALSE)
q4 <-q4[ , colSums(is.na(q4)) == 0]
```

## Our data chosen
      Column descriptions

**general info**      
  LIBNAME: Name of library
  STABR: State
  C_LEGBAS: Legal Basis Code
CC–City/County
CI–Municipal Government (city, town, or village)
CO–County/Parish
LD–Library District
MJ–Multi-jurisdictional
NL–Native American Tribal Government
NP–Non-profit Association or Agency
SD–School District
OT–Other
  POPU_LSA: Population of the legal service area (population group that the library caters to)

**locations & employees**
  CENTLIB: Number of central libraries
  BRANLIB: Number of branch libraries
  BKMOB: Number of bookmobiles
  MASTER: Number of full-time librarians with masters degrees in library science
  LIBRARIA: Number of full-time librarians.

**revenue**
  TOTINCM: *Total operating revenue* (DO NOT include revenue for major capital expenditures, contributions to endowments, revenue passed through to another agency (e.g., fines), or funds unspent in the previous fiscal year (e.g., carryover)). This is revenue given by federal, state, & local governments, monetary gifts and donations received in the current year, interest, library fines, fees for library services, etc.
  CAP_REV: *Total capital revenue* (revenue to be used for fixed cost items (funds
received for (a) site acquisition; (b) new buildings; (c) additions to or renovation        of library buildings; (d) furnishings, equipment, and initial collections (print,           nonprint, and electronic) for new buildings, building additions, or building                renovations; (e) computer hardware and software used to support library operations, to      link to networks, or to run information products; (f) new vehicles; and (g) other           one-time major projects.))

**operating expenditures**
  STAFFEXP: Total operating salaries & benefits cost for employees
  PRMATEXP: Total operating cost for print materials
  ELMATEXP: Total operating cost for electronic materials
  OTHMATEX: Total operating cost for other materials
  OTHOPEXP: Total operating cost for other expenditures not included in salaries, benefits, or materials
  
**capital expenditures**
  CAPITAL: Total capital expenditures

**collection**
  BKVOL: Print book collection (including duplicates)
  EBOOK: EBook licencing collection (include duplicates)
  VIDEO_PH: Physical video units (including duplicates)
  AUDIO_PH: Physical audio units (including duplicates)
  VIDEO_DL: Downloadable video units (including duplicates)
  AUDIO_DL: Downloadable audio units (including duplicates)
  OTHPHYS: Physical items aside from books, video, & audio (art supplies, cooking supplies, tools, etc.)
  
**metrics**
  VISITS: total persons entering the library per year (estimated or counted)
  REGBOR: total registered users
  ELMATCIR: Total use of Electronic materials
  PHYSCIR: Total use of Physical materials
  KIDCIRCL: Total use of Childrens materials, electronic or not.
  PITUSR: Total use of internet-connected computers
  WEBVISIT: Total times the website of the library has been visited
  
**inter-library relations**
  LOANTO: quantity of library materials provided to other libraries
  LOANFM: quantity of library materials recieved from other libraries
  
**synchronous program sessions**
  K0_5PRO: Number of Synchronous Program Sessions Targeted at Children Ages 0-5
  K6_11PRO: Number of Synchronous Program Sessions Targeted at Children Ages 6-11
  YAPRO: Number of Synchronous Program Sessions Targeted at Children Ages 12-18
  ADULTPRO: Number of Synchronous Program Sessions Targeted at Children Ages 19 and older
  TOTATTEN: Total attendance at Synchronous Programs

```{r selectingourdata}
ourdf <- df %>% select(LIBNAME, STABR, C_LEGBAS, POPU_LSA, CENTLIB, BRANLIB, BKMOB, MASTER, LIBRARIA, TOTINCM, CAP_REV, STAFFEXP, PRMATEXP, ELMATEXP, OTHMATEX, OTHOPEXP, CAPITAL, BKVOL, EBOOK, VIDEO_PH, AUDIO_PH, VIDEO_DL, AUDIO_DL, OTHPHYS, VISITS, REGBOR, ELMATCIR, PHYSCIR, KIDCIRCL, PITUSR, WEBVISIT, LOANTO, LOANFM, K0_5PRO, K6_11PRO, YAPRO, ADULTPRO, TOTATTEN)
```

## Our data agressively cleaned
This section takes out missing values present in the data in order to make sure the results are reasonable & accurate. 
Ncols: 192 -> 38
Nrow: 9248 -> 4955
```{r cleaningdata}
ourdfcleaned <- ourdf %>%
  filter(if_all(everything(), ~ !(. %in% c(-1, -3, -9))))
```

## Identifying problem columns
This section determines which columns have the most missing values so that we can determine if removing these columns is worth or not worth losing data points that don't have these values.

WEBVISIT & OTHPHYS will be removed because they had reasonably high levels of missing values and we had medium to low interest in these metrics. 

OTHOPEXP and STAFFEXP will be removed because they had significantly high levels of missing values even though we had reasonable interest in these metrics

K0_5PRO, K6_11PRO, YAPRO, ADULTPRO, and TOTATTEN will be removed because about 5% of libraries do not have these synchronous programs. Further research could be done comparing libraries that have these programs with libraries that dont to determine their effect, but for this research, ~500 library listings is too many to lose. 
```{r problemcols}
counts <- numeric(ncol(ourdf))

for (i in 1:ncol(ourdf)) {
  counts[i] <- sum(ourdf[[i]] %in% c(-1, -3, -9))
}

columnslevelofdifficulty <- data.frame(Column = colnames(ourdf), Count = counts)
columnslevelofdifficulty <- columnslevelofdifficulty %>% arrange(desc(Count))
head(columnslevelofdifficulty, n=12)
```

## Remaking a cleaned dataframe

The new dataframe created does not include K0_5PRO, K6_11PRO, YAPRO, ADULTPRO, TOTATTEN, WEBVISIT, OTHPHYS, OTHOPEXP, and STAFFEXP.

Ncols: 192 -> 38 -> 29
Nrow: 9248 -> 4955 -> 9161
```{r}
ourdf2 <- ourdf %>% select(-c(K0_5PRO, K6_11PRO, YAPRO, ADULTPRO, TOTATTEN, WEBVISIT, OTHPHYS, OTHOPEXP, STAFFEXP))

ourdfcleaned2 <- ourdf2 %>%
  filter(!if_any(everything(), ~ . %in% c(-1, -3, -9)))
```

## Exporting as CSV
```{r}
write.csv(ourdfcleaned2,"CleanedDf.csv")
```

